Job mobility and earnings growth: diffrerent markets, different outcomes.

### 2018-09-18 draft.Rmd
Update draft.Rmd

Worker mobility is a key strategy for improving earnings and conditions. Routinely, authors show how workers can make significant gains through quitting or securing promotions. Yet, capitalist countries differ in mobility rates; especially in terms of voluntary mobility. If quitting for new employer is more common in the UK than Germany, does this mean that German workers are held back from advancing their careers, or does it mean that the nature of quitting is different in both countries? Authors typically rationalise differences in job mobility as a difference in the freedom to hire or fire workers (Brown) or a difference in job creation and destruction (OECD). This article proposes that German and British institutions shape not only the likelihood of job mobility, but also the rewards tied to mobility.

The aim of the article is two-fold. First, we explicitly compare internal and external mobility as a means to improve earnings growth. Second, we compare this relationship under two different institutional settings, coordinated Germany and the liberal UK (Hall and Soskice, Thelen). Both aims stem from two wrinkles often found in the job mobility literature. First, authors' often ignore internal promotion as a form of voluntary mobility. Internal promotion is the main form of mobility desired by workers, yet we often focus on external quits. Second, is the over-reliance on single case studies. Authors exploring the impact of mobility often use single country exmaples despite the fact that european labour markets are highly divided. Our article highlights this nuance by explicitly comparing quits and promotions in two different institutional settings.

The theoretical literature has little to say on country differences in mobility rates, although certain hypotheses can be drawn out. In the *job-match* model, mobility stems from a mismatch between a worker's skills and her rewards or conditions (Jovanvich). In this way countries with low rates of job mobility, are simply countries with a good record of job matching. This would make sense in both Germany and the UK, where occupational licensing is prevalent in Germany, and less so in the UK. In the *job-search* model, mobility stems from the opportunity of higher wages. Thus if labour markets have a wide dispersion of wages, workers have more opportunity to match their reservation wages, which leads to more mobility. This too makes sense in the context of Germany and the UK. In gGermany wages inequality is far lower than in the UK, thus mobility between positions is less likely to lead to "wage-quits", when compared to the UK.  In *Human capital* theory, job mobility depends on the type of skills available to workers, either tranferable or firm specific. This approach may also apply to Germany and the UK, where German employers offer training in firm specific skills, while UK workers rely most on industry specific skills and tenure.

>>>>

Worker mobility is a key strategy for improving earnings and conditions. Routinely, authors show how workers make significant gains through quitting or securing promotions. Yet, capitalist countries differ in mobility rates; especially in terms of voluntary mobility. If quitting for a new employer is more common in the UK than Germany, does this mean that German workers are held back from advancing their careers, or does it mean that the nature of quitting is different in both countries? Authors typically rationalise country differences in mobility rates as differences in employment protection legislation (Brown) or differences in job creation and destruction (OECD). This article takes a different approach, and argues that German and British institutions shape not only the likelihood of job mobility, but also the rewards tied to mobility.

The aim of the article is two-fold. First, we explicitly compare internal and external mobility as a means of improving earnings growth. Second, we compare the relationship between mobility and earnings in two institutional settings, coordinated Germany and the liberal UK (Hall and Soskice, Thelen). Both aims stem from wrinkles in the job mobility literature. First, authors rarely see internal promotion as a form of voluntary mobility; instead they tend to focus on the effects of external quits alone. Second, authors exploring the impact of mobility often use single country exmaples despite the fact that european labour markets are highly divided. Our article highlights this nuance by explicitly comparing quits and promotions in two different institutional settings.

What can existing theories tell us about country differences in job mobility? In the *job-match* approach, found in both economic and sociological works, mobility stems from a mismatch between a worker's skills and her rewards or conditions (Jovanvich; Sorensen). Here, countries with low mobility are those capable of closely matching workers to jobs. By the same token, countries with high mobility, are those where job-matching is the responsibility of the worker. Since Germany is famous for its strong occupational licensing and its close links between skills gained in education and skills needed in the workplace, workers are better able to match their expectations with their resources quickly. Since the UK's occupational licensing is far weaker, and its links between educationa and workplace skills are also weaker, workers will need to quit and seek promotion several times before securing a "good fit". By the same toke, in the *job-search* model mobility stems from the opportunity of higher wages. Thus, if labour markets have a wide dispersion of wages and wage inequality, workers have more opportunity to match a "reservation wage" to a new position, which leads to higher rates of mobility. As a result, Germany mobility is low because wage inequality and wage dispersion is also low; mobility between positions is less likely to lead to "wage-quits" since differences between positions are minor. In the UK, mobility is high because wage dispersion or wage inequality between positions is also high; in this way a worker has more opportunities to find her "reservation wage".
